---
title: "Capstone Milestone Report"
author: "Warren T Ferrell Jewell"
date: "November 16, 2016"
output:
  html_document: default
  pdf_document: default
---

## About the Data
Data is provided by swiftkey in conjunction with Coursera Data Science CAPSTONE 
presented by Johns Hopkins University. The three English datasets were used for
this analysis.

* Twitter is 159 MB with 2360148 lines
* Blogs is   200 MB with 899288 lines
* News is    196 MB with 77259 lines

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, cache=TRUE, message=FALSE, warning=FALSE)
library(tm); library(ggplot2); library(dplyr)
dataPath <- "data/en_US/"
twitDir =  paste0(dataPath, "twitter/")
blogDir = paste0(dataPath, "blogs/")
newsDir = paste0(dataPath, "news/")
twitFile = paste0(twitDir, "en_US.twitter.txt")
blogFile = paste0(blogDir, "en_US.blogs.txt")
newsFile = paste0(newsDir, "en_US.news.txt")
twitLines = 2360148
blogLines = 899288
newsLines = 77259
twitChars = 162384825
blogChars = 370746263
newsChars = 386430028
twitSize = 159
blogSize = 200
newsSize = 196
readPartial <- function(filePath, NChars, xChars = NChars ) {
    fileCon <- file(filePath, "rt")
    partial <- list()
    if( xChars == "all")
        p = 1
    else
        p = xChars / NChars #probability that a line should be grabbed
    nGrabbed = 1
    while(length(line <- readLines(fileCon, n = 1L, warn=FALSE)) > 0) {
        if( runif(1) < p ) {
            x <- line %>% gsub(pattern = "#\\w+", replace = "") %>%
                gsub(pattern = "[^a-zA-Z '’.!,;]", replace = " ") %>% 
                gsub(pattern = "([[:punct:]])[[:punct:]]+", replace = "\\1") %>%
                gsub(pattern = "([[:punct:]])", replace = " \\1 ") # add space before/after punct
            #remove hashtags, nonenglish symbols, and repeated punctuation 
            partial <- c(partial, x) 
            nGrabbed = nGrabbed + 1
        }
    }
    close(fileCon)
    partialCorpus <- tm::VCorpus(VectorSource(partial),
        readerControl = list(reader = readPlain, language = "en_US"))
    meta(partialCorpus, tag = "file") <- basename(filePath)
    source <- gsub(filePath, pattern = paste0(dataPath,"([[:alpha:]]+)/.*"), replace = "\\1")
    meta(partialCorpus, tag = "source") <- source
    print(paste("Grabbed", nGrabbed, "lines from", source))
    return(partialCorpus)
}
getTDMstats <- function(x) {
    # where x is a TermDocumentMatrix
    list(sparsity = ifelse(!prod(dim(x)), 100, round((1 - length(x$v)/prod(dim(x))) * 100)) / 100,
         maxtermlength = max(nchar(Terms(x), type = "chars"), 0), 
         weightingLong = attr(x, "weighting")[1], 
         weightingShort = attr(x, "weighting")[2], 
         nonsparse = length(x$v), 
         sparse = prod(dim(x)) - length(x$v))
}

```

##Loading Data
Data is too big to work with all at once so a subset that is equal to about 10 MB
was used. Randomly selecting lines from the data. The tm package is slow to manipulate 
whole corpora (collection of text data) so some data cleaning was reformed during
data loading. These inclue removing non english characters, double punctuation,
and twitter hashtags. A space was also added before and after each punctuation mark
so they can be treated as part of an n-gram.

```{r loadingSubsetData, cache=TRUE, echo=FALSE}
xChars = 1E8 # work with about 100 MBs of data from each set
fileName <- paste0("RawCorpora", xChars, "characters")
if ( file.exists(fileName) ) {
    load(file = fileName)
} else {
    set.seed(4684)
    twit <- readPartial(twitFile, twitChars , xChars)  
    blog <- readPartial(blogFile, blogChars , xChars)
    news <- readPartial(newsFile, newsChars , xChars)
    corpora <- list(twit=twit,blog=blog, news=news)
    save(corpora, file = fileName)
}
rm(list = ls(pattern = "^[^c].*$")) #remove everything except corpora
```

## Basic look at data
To examine data we generate a term document matrix (TDM) from each corpus after
removing whitespace, and changing letters to all lower case. Punctuation and stopwords are left in because they could heavily affect text prediction. We ignore all terms that only occur once.

```{r clningData, cache=TRUE}
transformations <- list( content_transformer(tolower), stripWhitespace )
corpora <- lapply(corpora, function(x) 
            tm_map(x, tm_reduce, tmFuns = transformations))
tdm <- lapply(corpora, function(x) tm::TermDocumentMatrix(x, list(global = list(c(2,Inf)))))
tdm
```

## Cleaning Data
We see that each corpus still has thousands of terms and each term is extremely sparse.
A term Frequency vector cannot be generated on most computers by R's default as.matrix
function (can't allocate a 62 GB vector). Interestingly, news posts have the best
convergence with only 11551 terms but its sparsity (appears in few documents)
is still approximately 100%. This is due to a large number of 
terms that still only occur a few number of times - e.g. 11960 terms only appear once 
in the twitter corpora. To fix this we remove all terms with sparsity less than
0.999 from the TDM. This number was found experimentally. Sparsity of .99 produced only 128 terms from
twitter. An average native speaker has a vocabulary of 20,000 - 35,000 so a 128 term
model would likely not be able to predict even 1% of any 2 - grams they might devise. 
0.999 produced 1132, 3172, and 4830 terms respectively

```{r sparseTerms}
tdm.cln <- lapply(tdm, function(x) tm::removeSparseTerms(x, 0.999))
tdm.cln
termFreq <- lapply(tdm.cln, function(x) as.matrix(slam::rollup(x, 2, FUN = sum)))
termFreq <- lapply( termFreq, function(x) x[order(x, decreasing = TRUE), ])
lapply(termFreq, head)
lapply(termFreq, tail)
```

## Exploring Data Terms
This gets us down to a maximum term count of around 3000 which is a 
managable 37 billion searchspace for a 3-gram prediction algorithm. Using ggplot2
we graph the most common and least common terms from each dataset

```{r terms, echo=FALSE}
terms <- lapply(termFreq, FUN = function(x) data.frame(t(x)))
terms <- Reduce(function(x,y) merge(x,y, all = TRUE), terms)
row.names(terms) <- rev(names(tdm))
terms <- terms[ , apply(terms, 2, function(x) !anyNA(x))]
terms["source"] <- row.names(terms)
firstCols <- c(names(terms)[1:10], "source")
reshape::melt.data.frame( terms[,firstCols], id.vars = "source",
            value_name = "freq", variable_name = "term") %>%
    ggplot2::ggplot(aes(x = term, y = value, fill = source)) + 
        geom_bar( position = "dodge", stat = "identity", alpha = .3 ) +
        ylab("frequency") + ggtitle("Top terms")
lastCols <- c(rev(names(terms))[2:11], "source")
reshape::melt.data.frame( terms[,lastCols], id.vars = "source",
            value_name = "freq", variable_name = "term") %>%
    ggplot2::ggplot(aes(x = term, y = value, fill = source)) + 
        geom_bar( position = "dodge", stat = "identity", alpha = .3 ) +
        ylab("frequency") + ggtitle("Bottom terms")
```

## Remove infrequent terms from corpora 
Using the cleaned tdm we come up with a list of terms to keep from each corpus. 
Removing all words we don't want would be far more computationally 
expensive and the tm package provides a 
removeWords function but it works by forming a very large regex
expression. A regex that long is not supported by the compiler so we would need to use a 
loop to remove each term from the corpurs ourself.

```{r infrequentTerms}
commonTerms <- lapply(tdm.cln, function(x) x$dimnames$Terms)
for (i in seq_along(corpora)) {    # no easy pass by reference in R
    takeCommonTerms <- function(x) { 
        s <- strsplit(x, split = " ")
        y <- c()
        for(word in unlist(s)) {
            if(word %in% commonTerms[[i]]) 
                y <- paste(y, word, sep = " ")
        }
        return(y)
    }
    corpora[[i]] <- tm_map(corpora[[i]], content_transformer(takeCommonTerms))
}

```

## N-grams
Now that we have the most common terms we create all the potenial 2-4 grams 
that we will attempt to predict.

```{r makeNGrams, echo=FALSE}
ngram_tokenizer <- function(s, k = 2:4) {
        x <- strsplit(as.String(s), " ")[[1]]
        tokens <- c()
        for( n in k )
            tokens <- c(tokens, ngrams( x[(x != "") & (x != " ") ], n ))
        return( tokens )
}
ngram_tokenizer <- Token_Tokenizer(ngram_tokenizer)
cntrl <- list(tokenizer = ngram_tokenizer, global = list(c(2,Inf)))
gdm <- lapply(corpora, function(x) tm::TermDocumentMatrix(x, cntrl))
gramFreq <- lapply(gdm, function(x) as.matrix(slam::rollup(x, 2, FUN = sum)))     
gramFreq <- lapply(gramFreq, function(x) x[order(x, decreasing = TRUE), ])
gdm
lapply(gramFreq, head) #most frequent grams
lapply(gramFreq, tail) #least frequent grams
```

## Clean Ngrams
Too many grams to deal with so we must remove sparse ngrams. Then look at the bottom
ngrams to see if they are sensicle.

```{r cleanNgrams}
gdm.cln <- gdm
gdm.cln <- lapply(gdm, function(x) tm::removeSparseTerms(x, 0.999))
gramFreq <- lapply(gdm.cln, function(x) as.matrix(slam::rollup(x, 2, FUN = sum)))
gramFreq <- lapply(gramFreq, function(x) x[order(x, decreasing = TRUE), ])
gdm.cln
lapply(gramFreq, tail)
```

##Top Ngrams
Graph of top and bottom grams

```{r exploreGrams, echo = FALSE}
grams <- lapply(gramFreq, FUN = function(x) data.frame(t(x)))
grams <- Reduce(function(x,y) merge(x,y, all = TRUE), grams)
row.names(grams) <- rev(names(tdm))
grams <- grams[ , apply(grams, 2, function(x) !anyNA(x))]
grams["source"] <- row.names(grams)
firstCols <- c(names(grams)[1:3], "source")
reshape::melt.data.frame( grams[,firstCols], id.vars = "source",
                        value_name = "freq", variable_name = "gram") %>%
    ggplot2::ggplot(aes(x = gram, y = value, fill = source)) + 
        geom_bar( position = "dodge", stat = "identity", alpha = .3 ) +
        ylab("frequency") + ggtitle("Top grams")
lastCols <- c(rev(names(grams))[2:4], "source")
reshape::melt.data.frame( grams[,lastCols], id.vars = "source",
                value_name = "freq", variable_name = "gram") %>%
    ggplot2::ggplot(aes(x = gram, y = value, fill = source)) + 
        geom_bar( position = "dodge", stat = "identity", alpha = .3 ) +
        ylab("frequency") + ggtitle("Bottom grams")
```


## Model
Generate models based on the created Ngrams. Use a list of linked lists (essentially 
a hash table) to dramatically reduce the size of the model. Each term has a list
that points to terms that have been seen following that term. The final term in the
gram is stored in the list with a frequency variable that is used to rank which
is the most likely gram during prediction.

```{r model}
nGramModel.create <- function(gramFreq) {
    model <- structure(list(), class = "nGramModel")
    for (i in seq_along(gramFreq)) {
        gram <- eval(parse(text = names(gramFreq[i])))
        freq <- gramFreq[[i]]
        recurse <- function(model, gram, freq) {
            term = gram[[1]]
            if( length(gram) == 1 ) {
                model[[term]] <- list("gfreq" = freq)
                return( model )
            }
            else {
                model[[term]] <- recurse(model[[term]], gram[-1], freq)
                return( model )
            }
        }
        model  <- recurse(model, gram, freq) 
    }
    return ( model )
}
gramModel <- lapply(gramFreq, function(x) nGramModel.create(x))
save(gramModel, file = paste0(object.size(gramModel), "gramModel.RData"))
```

## Predict
With our generated model we predict some common phrases and see how the top 3 
suggestions vary across the datasets. The 

```{r predict, echo=FALSE}
nGramModel.predict <- function(model, pGram = list("")) {
    pGram <- tolower(pGram)
    if( is.character(pGram) & length(pGram) == 1 ){
        pGram <- as.character(strsplit(pGram, split = " ")[[1]])
    }
    if( length(pGram) > 3)
            pGram <- pGram[ (length(pGram)-2):length(pGram)] # Take only the last 3 terms
    if( length(pGram) == 0 ) {
        freqSort <- c(0)
        termSort <- c("")
        
        for( l in seq_along(model) ){
            freq <- 1
            term <- ""
            if( ! is.numeric(model[[l]]) ){
                term <- names(model)[[l]]
                if( is.null((freq <- model[[l]]$gfreq)))
                    freq <- 1
            } 
            for( t in seq_along(termSort) ) {
                if( freq > freqSort[[t]] ) {
                    freqSort <- append(freqSort, freq, after = t - 1)
                    termSort <- append(termSort, term, after = t - 1)
                    break # Found position
                } 
            }
        }
        return( termSort[1:3] )
    } else if ( pGram[[1]] %in% names(model) ) {
        pterm <- pGram[[1]]
        
        return( nGramModel.predict(model[[pterm]], pGram[-1]) )
    } else {
        return( nGramModel.predict(model, pGram[-1]) ) 
    }
}
lapply(gramModel, function(x) suppressWarnings(nGramModel.predict(x, "the")))
lapply(gramModel, function(x) suppressWarnings(nGramModel.predict(x, "thanks for")))
lapply(gramModel, function(x) suppressWarnings(nGramModel.predict(x, "you")))

```


## Thoughts
Need to remove grams that resulted from removing uncommon terms from the dataset. 


## references
https://eight2late.wordpress.com/2015/05/27/a-gentle-introduction-to-text-mining-using-r/

http://stackoverflow.com/questions/13614399/how-can-i-use-attr-with-lapply

http://stackoverflow.com/questions/32997201/how-to-store-sparsity-and-maximum-term-length-of-a-term-document-matrix-from-tm

http://www.economist.com/blogs/johnson/2013/05/vocabulary-size
