---
title: "Capstone Milestone Report"
author: "Warren T Ferrell Jewell"
date: "November 16, 2016"
output:
  html_document: default
  pdf_document: default
---

## About the Data

```{r setup, include=FALSE}
mainPath <- "C:/Users/warre/Dropbox/GitHub/CAPSTONE/"
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message=FALSE, warning=FALSE)
library(tm); library(ggplot2); library(dplyr); library(compiler); library(pryr)
library(foreach); library(doParallel); library(microbenchmark); library(fastmatch)
source(paste0(mainPath, 'scripts/proftable.R'))
source(paste0(mainPath, "scripts/cleanLine.R"))
#cleanLine <- compiler::cmpfun( cleanLine )
xChars = 1E6 # work with subset of data from each set
trainPath <- paste0(mainPath, "data/en_US/",xChars,"/")
cacheObj = FALSE
doParallel::registerDoParallel(cores= 4)
```

##Common Terms
Using on the top 20000 terms (minimum vocabulary for an adult)

```{r topTerms, echo = FALSE}
if ( file.exists("objects/termFreq") ) {
    load(file = "objects/termFreq")
} else {
    if ( file.exists("objects/tdm.trim") ) {
        load(file = "objects/tdm.trim")
    } else {
        source(paste0(mainPath, "scripts/tdm.trim.R"))
    }
    termFreq <- as.matrix(slam::rollup(tdm.trim, 2, FUN = sum))
    termFreq <- termFreq[order(termFreq, decreasing = TRUE), ]
    save(corpus, file = "objects/termFreq")
}

topTerms <- termFreq[1:20000]
includedTerms <- c(names(topTerms), "\"", ".", "?", "!", ",",";", ":", "&")
```


```{R oldLoading}
# source(paste0(mainPath, 'scripts/scripts/readSentence.R'))
# filePre <- paste0('objects/sentence/', xChars)
# readData <- compiler::cmpfun( readSentenceData )
# source(paste0(mainPath, 'scripts/readLines.R'))
# filePre <- paste0('objects/lines/', xChars)
# readData <- compiler::cmpfun( readLineData )
# source(paste0(mainPath, 'scripts/readWholeFile.R'))
# filePre <- paste0("objects/files/", xChars)
# keepTerms <- compiler::cmpfun( keepTerms )
# readData <- compiler::cmpfun( readFileData )
```

## Load Data



```{r loadData, echo=FALSE}
source(paste0(mainPath, 'scripts/references/readReferences.R'))
filePre <- paste0("objects/refs/", xChars)
readData <- compiler::cmpfun( readRefData )

fileName <- paste0(filePre, "corpus")
if ( cacheObj && file.exists(fileName) ) {
    load(file = fileName)
} else {
    Rprof("read.out")
    corpus <- readData(trainPath, includedTerms, xChars)
    Rprof(NULL)
    summaryRprof("read.out")
    save(corpus, file = fileName)
}
```


## N-grams
Now that we have all sentences

```{r makeNGrams, echo=FALSE}
source('C:/Users/warre/Dropbox/GitHub/CAPSTONE/scripts/gramTokenizer.R')
maxGram = 12
minFreq = 1
#formals(ngram_tokenizer)$k <- maxGram
ngram_tokenizer <- compiler::cmpfun( ngram_tokenizer )
cntrl <- list(tokenize = Token_Tokenizer(ngram_tokenizer),
              bounds = list(global = c(minFreq , Inf)))
fileName <- paste0(filePre, "gdm", "MinFreq", minFreq)
if ( file.exists(fileName) && cacheObj ) {
    load(file = fileName)
} else {
    gdm <- foreach(document = corpus, .combine = 'c',
                   .packages=c('magrittr','tm')) %dopar% {
        tm::VCorpus(VectorSource(document[[1]])) %>%
            tm::TermDocumentMatrix(control =  cntrl)
    }
    save(gdm, file = fileName)
}
gramFreq <- as.matrix(slam::rollup(gdm, 2, FUN = sum))
gramFreq <- gramFreq[order(gramFreq, decreasing = TRUE), ]
if( length(gramFreq) > 1e5 )
        gramFreq <- gramFreq[1:1e5]
gdm
head(gramFreq, n=6L) #most frequent grams
tail(gramFreq, n=6L) #least frequent grams
#cat(tail(names(gramFreq)), sep="\n")
```



## Tree
Generate trees by traversing each line in file. Use a list of linked lists (essentially 
a hash table) to dramatically reduce the size of the tree. Each term has a list
that points to terms that have been seen following that term. The final term in the
gram is stored in the list with a frequency variable that is used to rank which
is the most likely gram during prediction.


```{r corpusTree}
#rm(list=c("tdm.trim", "corpus","termFreq"))
# source('C:/Users/warre/Dropbox/GitHub/CAPSTONE/scripts/treeFromCorpus.R')
# treeConststr <- compiler::cmpfun( treeConststr )
# nGramTree.sort <- compiler::cmpfun( nGramTree.sort )
# nGramTree.create <- compiler::cmpfun( nGramTree.create )
```

```{r gdmTree}
#rm(list=c("tdm.trim", "corpus","termFreq"))
source('C:/Users/warre/Dropbox/GitHub/CAPSTONE/scripts/gdmTree.R')
treeConststr <- compiler::cmpfun( treeConststr )
gramOrder <- compiler::cmpfun( gramOrder )
nGramTree.create <- compiler::cmpfun( nGramTree.create )

```




```{r makeTree}
fileName <- paste0(filePre, "gramTree", "MinFreq", minFreq)
if ( file.exists(fileName) && cacheObj ) {
    load(file = fileName)
} else {
    gramTree <- nGramTree.create(gramFreq, 10)
    save(gramTree, file = fileName)
}

```



## Predict
With our generated tree we predict some common phrases and see how the top 3 
suggestions vary across the datasets. The 

```{r predictFunction, echo=FALSE}
source('C:/Users/warre/Dropbox/GitHub/CAPSTONE/scripts/predict.R')
treeSearch <- compiler::cmpfun( treeSearch )
formals(nGramTree.predict)$k <- maxGram
nGramTree.predict <- compiler::cmpfun( nGramTree.predict )

nGramTree.predict( gramTree, "in", maxGram )
nGramTree.predict( gramTree, "of", maxGram )
nGramTree.predict( gramTree, "case of", maxGram )
nGramTree.predict( gramTree, "a case of", maxGram )

```


```{r}
quiz <- c("The guy in front of me just bought a pound of bacon, a bouquet, and a case of", 
    "You're the reason why I smile everyday. Can you follow me please? It would mean the",
    "Hey sunshine, can you follow me and make me the",
    "Very early observations on the Bills game: Offense still struggling but the",
    "Go on a romantic date at the",
    "Well I'm pretty sure my granny has some old bagpipes in her garage I'll dust them off and be on my",
    "Ohhhhh #PointBreak is on tomorrow. Love that film and haven't seen it in quite some",
    "After the ice bucket challenge Louis will push his long wet hair out of his eyes with his little",
    "Be grateful for the good times and keep the faith during the",
    "If this isn't the cutest thing you've ever seen, then you must be")

```


```{r predict}

# lapply(quiz, function(x) nGramTree.predict( gramTree, x, maxGram ) )


```


## Thoughts
Need to remove grams that resulted from removing uncommon terms from the dataset. 


## references
https://eight2late.wordpress.com/2015/05/27/a-gentle-introduction-to-text-mining-using-r/

http://stackoverflow.com/questions/13614399/how-can-i-use-attr-with-lapply

http://stackoverflow.com/questions/32997201/how-to-store-sparsity-and-maximum-term-length-of-a-term-document-matrix-from-tm

http://www.economist.com/blogs/johnson/2013/05/vocabulary-size

http://stackoverflow.com/questions/31527345/how-to-add-unequal-length-named-vectors-in-r
